A Constraint Solving Approach to Parikh Images of Regular Languages

### 2024-02-22 evaluation.tex
A preliminary attempt at addressing the worse SAT performance
Adresses: #82

For our experiments, we modified the CE-Str solver to apply \Catra{}
 (with \Calculus{} as its chosen backend)
instead of the previous baseline method, resulting in a new back-end
\textbf{CA-Str}. Combining the results from SMT-COMP for \Ostrich{}~1.3 
with our new results running CA-Str we constructed a virtual 
portfolio~\Ostrich{}+CA that simulated running CA-Str as a fourth
back-end to \Ostrich{}.

@@

We extend the results of SMT-COMP~2023 with our modified \Ostrich{}
on the single-query string solving with linear integer arithmetic constraints
track (QF\_SLIA) in~\cref{tab:solve-status-smt-comp}. We obtain
the results by executing \Ostrich{}~1.3 with CA-Str using
the same benchmarking infrastructure (the StarExec cluster)
and configuration that ran SMT-COMP~2023, combining our new results for CA-Str with the
published results of SMT-COMP~2023.
\footnote{StarExec Job ID 59410, \enquote{SMT-COMP 2023 single query final}, 
available at \url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59410}.} 
Where avaiable, we use the revised, out-of
competition version of the results for solvers with bugs discovered during
the competition, including \Ostrich{} and Z3-Noodler. 
\footnote{StarExec Job ID 59668, \enquote{SMT-COMP 2023 single query final fixed} available at
\url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59668}.}
Note that Z3-Noodler abstained on 70~instances, and thus has a lower total number of results.
We used the parallel results for all solvers.

We picked QF\_SLIA since
\Ostrich{} already performed well on the other two string solving tracks. In fact,
every solver, including \Ostrich{}, handled every benchmark in QF\_SNIA within the timeout.
%
As can be seen from \cref{tab:solve-status-smt-comp} and \cref{fig:smt-comp:cactus},
integrating \Catra{} as a back-end
leads to gains both on satisfiable and unsatisfiable problems. On
satisfiable problems, the combination \Ostrich+CA is still
outperformed by cvc5 and z3alpha. On unsatisfiable benchmarks,
\Ostrich+CA now narrowly beats the other solvers 
(shown as another virtual portfolio solver, "Competitors") by squeezing past cvc5,
which is promising given that \Ostrich{}~1.3 (without CA-Str), cvc5, and z3alpha all show very strong
performance on this class of benchmarks. Similar improvements
occur in QF\_S, they are much smaller (an additional 2 unsatisfiable, 4 satisfiable 
solved out of~8\,847 in total).

>>>>

For our experiments, we modify the CE-Str solver to apply \Catra{}
 (with \Calculus{} as its chosen backend)
instead of the previous baseline method, resulting in a new back-end
\textbf{CA-Str}. Combining the results from SMT-COMP for \Ostrich{}~1.3 
with our new results running CA-Str we construct a virtual 
portfolio~\Ostrich{}+CA that simulates running CA-Str as a fourth
back-end to \Ostrich{}. We similarly combine the results of all non-\Ostrich{} solvers
to obtain the virtual portfolio solver~\textbf{Competition}.

@@

We extend the results of SMT-COMP~2023 with our modified \Ostrich{}
on the single-query string solving with linear integer arithmetic constraints
track (QF\_SLIA). We picked QF\_SLIA since
\Ostrich{} already performed well on the other two string solving tracks. In fact,
every solver, including \Ostrich{}, handled every benchmark in QF\_SNIA within the timeout.
Additionally, Parikh intersection problems would mainly be generated by \Ostrich{} when
solving constraints involving integers, meaning that CA-Str would be of little or no
help on the QF\_S track.

We obtain the results by executing \Ostrich{}~1.3 with CA-Str using
the same benchmarking infrastructure (the StarExec cluster)
and configuration that ran SMT-COMP~2023, combining our new results for CA-Str with the
published results of SMT-COMP~2023.
\footnote{StarExec Job ID 59410, \enquote{SMT-COMP 2023 single query final}, 
available at \url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59410}.} 
Where avaiable, we use the revised, out-of
competition version of the results for solvers with bugs discovered during
the competition, including \Ostrich{} and Z3-Noodler. 
\footnote{StarExec Job ID 59668, \enquote{SMT-COMP 2023 single query final fixed} available at
\url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59668}.}
Note that Z3-Noodler abstained on 70~instances, and thus has a lower total number of results.
We used the parallel results for all solvers.

As can be seen from \cref{tab:solve-status-smt-comp} and \cref{fig:smt-comp:cactus},
integrating \Catra{} as a back-end
leads to gains both on satisfiable and unsatisfiable problems. On
satisfiable problems, the combination \Ostrich+CA is still
outperformed by cvc5 and z3alpha. On unsatisfiable benchmarks,
\Ostrich+CA now narrowly beats the other solvers squeezing past cvc5,
which is promising given that \Ostrich{}~1.3 (without CA-Str), cvc5, and z3alpha all show very strong
performance on this class of benchmarks.

These results are expected. It is known that automata-based string solvers (\Ostrich{}, Z3-Noodler) 
tend to perform better on unsatisfiable than on satisfiable benchmarks, compared to solvers
that do not utilize automata and directly work on regular expressions (cvc5, z3alpha).
This is not unexpected, as the computation of automata representations of regular 
constraints can be expensive, and might be unnecessary for satisfiable formulas.
In addition, the algorithm in \Ostrich{} has stronger theoretical completeness
guarantees as cvc5 and Z3-Str, and ensuring completeness often has an adverse
effect on performance in practice~\cite{ostrich,Z3-str,cvc5}.
